---
title: "Inclassmod3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(palmerpenguins)
library(tidyverse)
```

# 1
Produce a scatterplot of the two variables. How would you describe the relationship
between the two variables? Be sure to label the axes and give an appropriate title.
Based on the appearance of the plot, does a simple linear regression appear reasonable
for the data?

```{r}
penguins %>% 
  ggplot(aes(body_mass_g, flipper_length_mm)) + 
  geom_point() +
  labs(title="Flipper Length vs Body Mass", x="Body Mass (g)", y="Flipper Length (mm)") + 
  theme_bw()
```

# 2

Produce a similar scatterplot, but with different colored plots for each species. How
does this scatterplot influence your answer to the previous part?
```{r}
penguins %>% 
  ggplot(aes(body_mass_g, flipper_length_mm, color=species)) + 
  geom_point() +
  labs(title="Flipper Length vs Body Mass", x="Body Mass (g)", y="Flipper Length (mm)") + 
  theme_bw()
```

# 3
Regardless of your answer to the previous part, produce a scatterplot of body mass
and flipper length for Gentoo penguins. Based on the appearance of the plot, does a
simple linear regression appear reasonable for the data?


```{r}
penguins %>% 
  filter(species == "Gentoo") %>%
  ggplot(aes(flipper_length_mm,body_mass_g, color=species)) + 
  geom_point() +
  labs(title="Flipper Length vs Body Mass", x="Body Mass (g)", y="Flipper Length (mm)") + 
  theme_bw()
```

# 4
What is the correlation between body mass and flipper length for Gentoo penguins.
Interpret this correlation contextually. How reliable is this interpretation?

```{r}
cor(penguins[,c("flipper_length_mm", "body_mass_g")], use = "complete.obs")
```

# 5
Use the lm() function to fit a linear regression for body mass and flipper length for
Gentoo penguins. Write out the estimated linear regression equation.

```{r}
data = penguins %>%
  filter(species == "Gentoo")
result<-lm(body_mass_g~flipper_length_mm, data=data)
summary(result)
```
$\hat{y} = -6787.281 + 54.623x$

# 6
Interpret the estimated slope contextually.

# 7
Does the estimated intercept make sense contextually?

# 8
Report the value of $R^2$ from this linear regression, and interpret its value contextually.
0.759

# 9
What is the estimated value for the standard deviation of the error terms for this
regression model, $\hat{\sigma}$ ? 

# 9
For a Gentoo penguin which has a flipper length of 220mm, what is its predicted body
mass in grams?
```{r}
newData = data.frame(c(220))
colnames(newData) = "flipper_length_mm"
predict(result, newData)
```

# 10
Produce the ANOVA table for this linear regression. Using only this table, calculate
the value of $R^2$.

```{r}
anova(result)
anova.tab = anova(result)
SST<-sum(anova.tab$"Sum Sq")
anova.tab$"Sum Sq"[1]/SST
```



