---
title: "Inclass"
author: "Hyun Suk (Max) Ryoo (hr2ee)"
date: "10/5/2021"
output: pdf_document
---

```{r}
library(dplyr)
library(MASS)
library(tidyverse)
library(faraway)
setwd("/Users/maxryoo/Documents/MSDS/STAT6021/Module7")
data <- seatpos
head(data)
```


## 1) Fit the full model with all the predictors. Using the summary() function, comment on
the results of the t tests and ANOVA F test from the output.
```{r}
result <- lm(hipcenter~., data=data)
summary(result)
```

This model is good for predicting hte hipcenter however, each predictor has a very high p-value, which means it is insignicant. We need to do more testing to drop the predictors

## 2)  Briefly explain why, based on your output from part 1, you suspect the model shows
signs of multicollinearity.

This model is good for predicting hte hipcenter however, each predictor has a very high p-value, which means it is insignicant. We need to do more testing to drop the predictors


## 3) . Provide the output for all the pairwise correlations among the predictors. Comment
briefly on the pairwise correlations

```{r}
pairs(data, lower.panel = NULL)
cor(data)
```

Lots of correlations that high between the predictors. I.E. Weight, HtShoes, Ht, Seated .. All of the variables seem to be correlated with one noather to be frankly honest. 

### 4) Check the variance inflation factors (VIFs). What do these values indicate about
multicollinearity?

```{r}
vif(result)
```

## 5) Looking at the data, we may want to look at the correlations for the variables that
describe length of body parts: HtShoes, Ht, Seated, Arm, Thigh, and Leg. Comment
on the correlations of these six predictors

Highly Correlated

## 6) Since all the six predictors from the previous part are highly correlated, you may decide
to just use one of the predictors and remove the other five from the model. Decide
which predictor out of the six you want to keep, and briefly explain your choice.
Thigh because it has a lower vif value after arm SHOULD BE HEIGHT SINCE HIGHER VIF
```{r}
result2 <- lm(hipcenter~Thigh, data=data)
summary(result2)
```
## 7) Since all the six predictors from the previous part are highly correlated, you may decide
to just use one of the predictors and remove the other five from the model. Decide
which predictor out of the six you want to keep, and briefly explain your choice.
```{r}

result3 <- lm(hipcenter~Ht + Arm + Weight, data=data)
summary(result3)
vif(result3)
```

Should be okay

## 8) Conduct a partial F test to investigate if the predictors you dropped from the full
model were jointly insignificant. Be sure to state a relevant conclusion.

```{r}
anova(result3,result)
```



Produce a plot of residuals against fitted values for your model from part 7. Based on
the residual plot, comment on the assumptions for the multiple regression model. Also
produce an ACF plot and QQ plot of the residuals, and comment on the plots.

```{r}
yhat = result3$fitted.values
res = result3$residuals
data %>%
  ggplot(aes(yhat, res)) + geom_point() + theme_bw() + geom_hline(yintercept=0, color="red")

```

```{r}
acf(result3$residuals, main="ACF Plot of Residuals with ystar")

```

```{r}
{
qqnorm(result3$residuals)
qqline(result3$residuals, col="red")
}
```

Produce a plot of residuals against fitted values for your model from part 7. Based on
the residual plot, comment on the assumptions for the multiple regression model. Also
produce an ACF plot and QQ plot of the residuals, and comment on the plots.
```{r}
summary(result3)
```

```{r}
summary(result)

```





