---
title: "Hw12"
author: "Hyun Suk (Max) Ryoo (hr2ee)"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up 

```{r}
library(tidyverse)
library(palmerpenguins)
library(gridExtra)
library(ROCR)
Data<-penguins
##remove penguins with gender missing
Data<-Data[complete.cases(Data[ , 7]),-c(2,8)]
##80-20 split
set.seed(1)
sample<-sample.int(nrow(Data), floor(.80*nrow(Data)), replace = F)
train<-Data[sample, ]
test<-Data[-sample, ]
head(train)
```

## 1 - A
We first need to recreate the model.
```{r}
result<-glm(sex ~ . - flipper_length_mm, family="binomial", data=train)
summary(result)
```

```{r}
##predicted survival rate for test data based on training data
preds<-predict(result,newdata=test, type="response")
##transform the input data into a format that is suited for the
##performance() function
rates<-prediction(preds, test$sex)
##store the true positive and false positive rates
roc_result<-performance(rates,measure="tpr", x.measure="fpr")
##plot ROC curve and overlay the diagonal line for random guessing
plot(roc_result, main="ROC Curve for Penguins")
lines(x = c(0,1), y = c(0,1), col="red")
```

Since this ROC curve is above the diagonal line, the logistic regression performs better than
random guessing.

## 1 - B

```{r}
##compute the AUC
auc<-performance(rates, measure = "auc")
auc@y.values
```

The AUC of our ROC curve is 0.9214286, which measn our logitic regressino does better than random guessing.

## 1 - C
```{r}
table(test$sex, preds>0.5)
```

From the table above we can compute the false positive rate, false negative rate, and error rate.

* The false positive rate (FRP) is $\frac{7}{7+28} \rightarrow \frac{7}{35} \rightarrow 0.2$.

* The false negative rate (FNR) is $\frac{4}{4 + 28} \rightarrow \frac{4}{32} \rightarrow 0.125$

* The error rate is 1 - accuracy = $1 - \frac{28 + 28}{28+28+7+4} = 1 - \frac{56}{67} = 1 - 0.8358209 = 0.1641791$

## 1 - D

```{r}
test<-data.frame(test,preds)
ggplot(test,aes(x=preds))+
geom_density()+
labs(title="Density Plot of Predicted Probs")
```

I personally don't believe the threshold needs to be changed. Sometimes threshold needs to be changed depending on the context of the analysis being done, but in this case increasing and decreasing doesn't make a huge difference in context of the problem. Also the Density plot shows that there is no huge difference in the prediction (probabilities), which means that 0.5 is a decent threshold. 

