---
title: "Project"
author: "Hyun Suk (Max) Ryoo (hr2ee)"
date: "11/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data processing
```{r}
## Data Processing
library(tidyverse)
library(dplyr)
library(MASS)
library(leaps)
setwd("/Users/maxryoo/Documents/MSDS/STAT6021/Project2")
data <- read.csv("data/insurance.csv")
head(data)
```

```{r}
data$significant.charge = as.factor(data$charges > median(data$charges))
data$smoker = as.factor(data$smoker)
data$region = as.factor(data$region)
head(data)
```

## Initial EDA

```{r}
ggplot(aes(x=bmi, y=charges, color=smoker), data=data) + 
  labs(title="Scatter Plot of Charges vs BMI by Smoker Status") +
  theme_bw() +
  geom_point()
```

```{r}
ggplot(aes(x=age,y=charges, color=smoker, size=bmi), data=data) + 
  labs(title="Scatter plot of Charges vs Age by BMI and Smoker Status") +
  theme_bw() +
  geom_point(alpha=0.5)
```

```{r}
ggplot(aes(x=age,y=bmi, color=significant.charge), data=data) + 
  labs(title="Scatter plot of Charges vs Age by BMI and Smoker Status") +
  theme_bw() +
  geom_point(alpha=0.5)
```

```{r}
ggplot(data, aes(x=region, y=age, fill=as.factor(significant.charge)))+
  geom_boxplot() +
  theme_bw() +
  labs(x="region", y="ages", title="Dist of bmi by region and smoker status")
```

```{r}
ggplot(data, aes(x=as.factor(children), y=charges))+
  geom_boxplot() +
  theme_bw() +
  labs(x="region", y="ages", title="Dist of bmi by region and smoker status")
```

```{r}
ggplot(data, aes(x=smoker, y=bmi, fill=as.factor(significant.charge)))+
  geom_boxplot() +
  theme_bw() +
  labs(x="Region", y="bmi", title="Distribution of bmi by region and smoker status") +  scale_y_continuous(limits=c(10,60), breaks=seq(10,60,5)) 
```


## Correlation

```{r}
pairs(data[c("age", "bmi", "charges")])
```

```{r}
round(cor(data[c("age", "bmi", "charges")]),4)
```

## Blind Full Model
```{r}
mlr_full = lm(charges ~  age + bmi + children + smoker + region, data=data)
summary(mlr_full)
```
The full regression is as follows.

## Assumption Check of Full Model
```{r}
yhat_full <- mlr_full$fitted.values
res_full <- mlr_full$residuals
data %>%
  ggplot(aes(yhat_full, res_full)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

The residuals are obviously not evenly scattered, which then we can utilize the boxcox method to give us information about transformation.

```{r}
boxcox(mlr_full, lambda=seq(0,0.3, 0.01))
```

From the boxcox we can try a lambda value of 0.15 for transformation.

QQPlot Normality of Full Model

```{r}
{
  qqnorm(mlr_full$residuals)
  qqline(mlr_full$residuals, col="red")
}
```

ACF of full model

```{r}
acf(mlr_full$residuals, main="ACF")
```


## Transformation of full model
```{r}
first_transformation_full <- data
first_transformation_full$charges <- first_transformation_full$charges^0.15
mlr_transform_first <- lm(charges ~ age + bmi + children + smoker + region, data=first_transformation_full)
summary(mlr_transform_first)
```

Residual Plot of the transformed model.

```{r}
yhat_full_t1 <- mlr_transform_first$fitted.values
res_full_t1 <- mlr_transform_first$residuals
data %>%
  ggplot(aes(yhat_full_t1, res_full_t1)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

Violation in constant variance

```{r}
boxcox(mlr_transform_first, lambda=seq(0,3, 0.01))
```

QQPLOT of transformed full model.

```{r}
{
  qqnorm(mlr_transform_first$residuals)
  qqline(mlr_transform_first$residuals, col="red")
}
```

## Why is this happening? Is there some weird behavior in the response variable?

```{r}
hist(data$charges)
```

```{r}
hist(first_transformation_full$charges)
```


## Trial of other predictors to fullfill the linearity assumption.

Maybe we can add some interaction terms to the model to see if we can fix the linearity assumption.

```{r}
interaction_age_bmi_with_smoker = lm(charges ~  age*smoker + bmi*smoker + children  + region, data=data)
summary(interaction_age_bmi_with_smoker)
```

Residual Plot
```{r}
yhat_full_first_inter <- interaction_age_bmi_with_smoker$fitted.values
res_full_first_inter <- interaction_age_bmi_with_smoker$residuals
data %>%
  ggplot(aes(yhat_full_first_inter, res_full_first_inter)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

This residual plot is a little better, lets see if we can transform the response with this new equation.

```{r}
boxcox(interaction_age_bmi_with_smoker, lambda=seq(0,0.3, 0.01))
```

Maybe we can use a lambda value of 0.125

```{r}
interaction_transform <- data
interaction_transform$charges <- interaction_transform$charges^0.125
mlr_interaction_tranform <- lm(charges ~  age*smoker + bmi*smoker + children  + region, data=interaction_transform)
summary(mlr_interaction_tranform)
```

Recheck Residual Plot

```{r}
yhat_full_first_inter_tranform <- mlr_interaction_tranform$fitted.values
res_full_first_inter_tranform <- mlr_interaction_tranform$residuals
data %>%
  ggplot(aes(yhat_full_first_inter_tranform, res_full_first_inter_tranform)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

Still see the same without adding the interaction terms.

```{r}
boxcox(mlr_interaction_tranform, lambda=seq(0,3, 0.01))
```


Still no luck. We retried this many times, but weren't lucky.

## Partial F test of the interaction vs simple model after two transformation of response variable
```{r}
full <- mlr_interaction_tranform
reduced <- lm(charges ~ age + bmi + children + smoker + region, data=interaction_transform)
anova(reduced, full)
```

We can't drop the interaction terms.

## Try different models before splitting the data

```{r}
bic_selection_model = lm(charges ~  age + bmi + children  + smoker, data=data)
summary(bic_selection_model)
```

Residual Plot

```{r}
yhat_bic <- bic_selection_model$fitted.values
res_bic <- bic_selection_model$residuals
data %>%
  ggplot(aes(yhat_bic, res_bic)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

We see a similar plot. Transformation?

```{r}
boxcox(bic_selection_model, lambda=seq(0,0.3, 0.01))
```

Again lambda of 0.15

```{r}
bic_transform <- data
bic_transform$charges <- bic_transform$charges^(0.15)
bic_selection_model_transform = lm(charges ~  age + bmi + children  + smoker, data=bic_transform)
summary(bic_selection_model_transform)
```

Residual Plot

```{r}
yhat_bic_transform <- bic_selection_model_transform$fitted.values
res_bic_transform <- bic_selection_model_transform$residuals
data %>%
  ggplot(aes(yhat_bic_transform, res_bic_transform)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```


QQPlot
```{r}
{
  qqnorm(bic_selection_model_transform$residuals)
  qqline(bic_selection_model_transform$residuals, col="red")
}
```

# See the two groupings

## Check the data points below and above 20,000

```{r}
less_charge <- data[data$charges < 20000,]
more_change <- data[data$charges >= 20000, ]
library(tidyverse)
ggplot(aes(x=bmi, y=charges, color=smoker), data=less_charge) + 
  labs(title="Scatter Plot of Charges vs BMI by Smoker Status for charges < 20,000") +
  theme_bw() +
  geom_point()
ggplot(aes(x=age,y=charges, color=smoker), data=less_charge) + 
  labs(title="Scatter plot of Charges vs Age by Smoker Status charges < 20,000") +
  theme_bw() +
  geom_point(alpha=0.5)

ggplot(aes(x=bmi, y=charges, color=smoker), data=more_change) + 
  labs(title="Scatter Plot of Charges vs BMI by Smoker Status charges > 20,000") +
  theme_bw() +
  geom_point()

ggplot(aes(x=age,y=charges, color=smoker), data=more_change) + 
  labs(title="Scatter plot of Charges vs Age by Smoker Status charges > 20,000") +
  theme_bw() +
  geom_point(alpha=0.5)
```

```{r}
pairs(more_change[c("age", "bmi", "charges")])
pairs(less_charge[c("age", "bmi", "charges")])
```

## Split dataset by smoking status

```{r}
smokers <- data[data$smoker == 'yes',]
non_smokers <- data[data$smoker != 'yes',]

ggplot(aes(x=bmi, y=charges, color=smoker), data=smokers) + 
  labs(title="Scatter Plot of Charges vs BMI For Smokers") +
  theme_bw() +
  geom_point()
ggplot(aes(x=age,y=charges, color=smoker), data=smokers) + 
  labs(title="Scatter plot of Charges vs Age For Smokers") +
  theme_bw() +
  geom_point(alpha=0.5)

ggplot(aes(x=bmi, y=charges, color=smoker), data=non_smokers) + 
  labs(title="Scatter Plot of Charges vs BMI For Non_Smokers") +
  theme_bw() +
  geom_point()

ggplot(aes(x=age,y=charges, color=smoker), data=non_smokers) + 
  labs(title="Scatter plot of Charges vs Age For Non_Smokers") +
  theme_bw() +
  geom_point(alpha=0.5)
```


## Smokers forward selection

```{r}
## Smokers
##intercept only model
regnull_smoker <- lm(charges~1, data=smokers)
##model with all predictors
regfull_smoker <- lm(charges ~ age + sex + bmi + children + region , data=smokers)
```

Forward Selection
```{r}
step(regnull_smoker, scope=list(lower=regnull_smoker, upper=regfull_smoker), direction="forward")
```


## Non Smokers forward selection
```{r}
#Non smokers
regnull_non_smokers <- lm(charges~1, data=non_smokers)
##model with all predictors
regfull_non_smokers <- lm(charges ~ age + sex + bmi + children + region , data=non_smokers)
```

Forward Selection
```{r}
step(regnull_non_smokers, scope=list(lower=regnull_non_smokers, upper=regfull_non_smokers), direction="forward")
```

## We get different models (first smokers)

```{r}
mlr_full_smoker = lm(charges ~  bmi+age , data=smokers)
summary(mlr_full_smoker)
```

Residual Plot of smokers looks the best so far.

```{r}
yhat_full_smoker <- mlr_full_smoker$fitted.values
res_full_smoker <- mlr_full_smoker$residuals
smokers %>%
  ggplot(aes(yhat_full_smoker, res_full_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

```{r}
boxcox(mlr_full_smoker)
```

Rseponse needs to be transformed 

```{r}
smokers_transform <- smokers
smokers_transform$charges <- smokers_transform$charges^0.5
mlr_full_smoker_transform = lm(charges ~  bmi+age , data=smokers_transform)
mlr_full_smoker_transform_full = lm(charges ~  bmi+age + sex + region , data=smokers_transform)
summary(mlr_full_smoker_transform)
```

residual plot for transformed response variable for smoerks

```{r}
yhat_full_smoker <- mlr_full_smoker_transform$fitted.values
res_full_smoker <- mlr_full_smoker_transform$residuals
smokers %>%
  ggplot(aes(yhat_full_smoker, res_full_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

```{r}
boxcox(mlr_full_smoker_transform)
```

QQPlot Smokers

```{r}
{
  qqnorm(mlr_full_smoker_transform$residuals)
  qqline(mlr_full_smoker_transform$residuals, col="red")
}
```

ACF for somkers

```{r}
acf(mlr_full_smoker_transform$residuals, main="ACF")
```

## Can we drop the other predictors?
```{r}
mlr_full_smoker_full = lm(charges ~  bmi+age+children+region , data=smokers_transform)
summary(mlr_full_smoker_full)
```


## Residual of full model for smokers
```{r}
yhat_full_smoker <- mlr_full_smoker_full$fitted.values
res_full_smoker <- mlr_full_smoker_full$residuals
smokers %>%
  ggplot(aes(yhat_full_smoker, res_full_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

Residual Seems Fine, can we drop these predictors using partial F - test.
```{r}
summary(mlr_full_smoker_transform)
anova(mlr_full_smoker_transform,mlr_full_smoker_transform_full )
```

We can drop these other predictors since p-value is above 0.05

## Outliers detection

```{r}
##critical value using Bonferroni procedure
n<-dim(smokers_transform)[1]
p<-3
crit<-qt(1-0.05/(2*n), n-1-p)
##externally studentized residuals
ext.student.res<-rstudent(mlr_full_smoker_transform)
##identify
ext.student.res[abs(ext.student.res)>crit]
```

```{r}
##leverages
lev<-lm.influence(mlr_full_smoker_transform)$hat
##identify
round(lev[lev>2*p/n],4)
```

```{r}
DFFITS<-dffits(mlr_full_smoker_transform)
round(DFFITS[abs(DFFITS)>2*sqrt(p/n)],3)
```

```{r}
DFBETAS<-dfbetas(mlr_full_smoker_transform)
tempdfbetas = abs(DFBETAS)>2/sqrt(n)
tempdfbetas[(tempdfbetas[,1] == TRUE | tempdfbetas[,2] == TRUE | tempdfbetas[,3] == TRUE ) ,]
```

```{r}
COOKS<-cooks.distance(mlr_full_smoker_transform)
COOKS[COOKS>qf(0.5,p,n-p)]
```


## Non Smokers Analysis

Predictors from forward selection process.

```{r}
mlr_full_non_smoker = lm(formula = charges ~ age + children + region + sex, data = non_smokers)
summary(mlr_full_non_smoker)
```

```{r}
yhat_full_non_smoker <- mlr_full_non_smoker$fitted.values
res_full_non_smoker <- mlr_full_non_smoker$residuals
non_smokers %>%
  ggplot(aes(yhat_full_non_smoker, res_full_non_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

Can we transform the response variable?

```{r}
boxcox(mlr_full_non_smoker, c(-0.5, 0.5, 0.1))
```

```{r}
non_smoker_transform <- non_smokers
non_smoker_transform$charges <- non_smoker_transform$charges^(-0.1)
mlr_full_non_smoker_transform = lm(formula = charges ~ age + children + region + sex, data = non_smoker_transform)
summary(mlr_full_non_smoker_transform)
```

Residual plot of transformed response variable with forward selection for non smokers

```{r}
yhat_full_non_smoker <- mlr_full_non_smoker_transform$fitted.values
res_full_non_smoker <- mlr_full_non_smoker_transform$residuals
non_smokers %>%
  ggplot(aes(yhat_full_non_smoker, res_full_non_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```


```{r}
boxcox(mlr_full_non_smoker_transform)
```

QQPlot

```{r}
{
  qqnorm(mlr_full_non_smoker_transform$residuals)
  qqline(mlr_full_non_smoker_transform$residuals, col="red")
}
```

ACF

```{r}
acf(mlr_full_smoker_transform$residuals, main="ACF")
```

## Dropping of outliers

We transformed and tried many steps, but we can see that the outliers are actually small sub populations. It could be that they are some clients with preeixsting conditions, if that is the case they will follow a different trend and will be classified as outliers for our multiple linear regression model. 

```{r}
non_smokers %>%
  ggplot(aes(x=age, y=charges, color=region, size=children)) + 
  theme_bw() +
  geom_point(alpha=0.5)
```


Lets drop these outliers using dffits

```{r}
drop_outleirs <- non_smokers
DFFITS_drop1<-dffits(mlr_full_non_smoker)
want_drop1 = names(DFFITS_drop1[abs(DFFITS_drop1)>2*sqrt(5/dim(non_smokers)[1])])
drop_outleirs = non_smokers[setdiff(rownames(non_smokers), want_drop1),]
percent_dropped = (dim(non_smokers)[1] - dim(drop_outleirs)[1]) / dim(non_smokers)[1]
percent_dropped*100
```

Make a multiple linear regression model after dropping the outliers

```{r}
mlr_full_non_smoker_transform_drop = lm(formula = charges ~  region + age + children + sex, data = drop_outleirs)
summary(mlr_full_non_smoker_transform_drop)
```


Residual plot of the model constructed after dropping the first set of outliers

```{r}
yhat_full_non_smoker <- mlr_full_non_smoker_transform_drop$fitted.values
res_full_non_smoker <- mlr_full_non_smoker_transform_drop$residuals
drop_outleirs %>%
  ggplot(aes(yhat_full_non_smoker, res_full_non_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

Double checking EDA to see if the top section was fully dropped.

```{r}
drop_outleirs %>%
  ggplot(aes(x=age, y=charges, color=region, size=children)) + 
  theme_bw() +
  geom_point(alpha=0.5)
```


Since not all points were dropped, lets try dropping again based on outliers for DFFITS.
```{r}
drop_outleirs2 <- drop_outleirs
DFFITS_drop2<-dffits(mlr_full_non_smoker_transform_drop)
want_drop2 = names(DFFITS_drop2[abs(DFFITS_drop2)>2*sqrt(5/dim(drop_outleirs)[1])])
drop_outleirs2 = drop_outleirs[setdiff(rownames(drop_outleirs), want_drop2),]
(dim(drop_outleirs)[1] - dim(drop_outleirs2)[1]) / dim(drop_outleirs)[1]
(dim(non_smokers)[1] - dim(drop_outleirs2)[1]) / dim(non_smokers)[1]
```

```{r}
drop_outleirs2 %>%
  ggplot(aes(x=age, y=charges, color=region, size=children)) + 
  geom_point(alpha = 0.4) + theme_bw()
```

```{r}
mlr_full_non_smoker_transform_drop2 = lm(formula = charges ~  region + age + children + sex, data = drop_outleirs2)
summary(mlr_full_non_smoker_transform_drop2)
```

```{r}
yhat_full_non_smoker <- mlr_full_non_smoker_transform_drop2$fitted.values
res_full_non_smoker <- mlr_full_non_smoker_transform_drop2$residuals
drop_outleirs2 %>%
  ggplot(aes(yhat_full_non_smoker, res_full_non_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```

```{r}
boxcox(mlr_full_non_smoker_transform_drop2)
```

```{r}
{
  qqnorm(mlr_full_non_smoker_transform_drop2$residuals)
  qqline(mlr_full_non_smoker_transform_drop2$residuals, col="red")
}
```

Final transformation

```{r}
mlr_full_non_smoker_transform_drop3 = lm(formula = (charges)^0.5 ~  region + age + children + sex, data = drop_outleirs2)
summary(mlr_full_non_smoker_transform_drop3)
```

Residual Plot
```{r}
yhat_full_non_smoker <- mlr_full_non_smoker_transform_drop3$fitted.values
res_full_non_smoker <- mlr_full_non_smoker_transform_drop3$residuals
drop_outleirs2 %>%
  ggplot(aes(yhat_full_non_smoker, res_full_non_smoker)) + 
  geom_point() + 
  theme_bw() +
  geom_hline(yintercept = 0, color="red")
```


QQ PLOT

```{r}
{
  qqnorm(mlr_full_non_smoker_transform_drop3$residuals)
  qqline(mlr_full_non_smoker_transform_drop3$residuals, col="red")
}
```



## Logistic
```{r}
set.seed(6021) ##for reproducibility
sample<-sample.int(nrow(data), floor(.70*nrow(data)), replace = F)
train<- data[sample, ] ##training data frame
test<-data[-sample, ] ##test data frame
result<-glm(significant.charge ~ age + bmi + children + smoker + region + sex, family="binomial", data=train)
summary(result)
```

```{r}
library(ROCR)
##predicted survival rate for test data based on training data
preds<-predict(result,newdata=test, type="response")
##transform the input data into a format that is suited for the
##performance() function
rates<-prediction(preds, test$significant.charge)
##store the true positive and false positive rates
roc_result<-performance(rates,measure="tpr", x.measure="fpr")
##plot ROC curve and overlay the diagonal line for random guessing
plot(roc_result, main="ROC Curve")
lines(x = c(0,1), y = c(0,1), col="red")
```


```{r}
##compute the AUC
auc<-performance(rates, measure = "auc")
auc@y.values
```


Matrix
```{r}
table(test$significant.charge, preds>0.5)
```

Threshold value manipulation
```{r}
table(test$significant.charge, preds>0.25)
```
Doesn't play a huge role in decreasing the False Positive Rate. We want to make sure that when someone signs up for a plan that they don't get charged significantly given their condition.

```{r}
test<-data.frame(test,preds)
ggplot(test,aes(x=preds))+
geom_density()+
labs(title="Density Plot of Predicted Probs") + theme_bw()
```