---
title: "In class"
author: "Hyun Suk (Max) Ryoo (hr2ee)"
date: "9/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(MASS)
```

# 1
Create a scatter plot of brain weight against body weight of land mammals. Comment
on the appearance of the plot. Do any assumptions for simple linear regression appear
to be violated? If so, which ones?
```{r}
mammals %>%
  ggplot(aes(body, brain)) + geom_point() + theme_bw()
```

## Fit a simple linear regression to the data, and create the corresponding residual plot.
Do any assumptions for simple linear regression appear to be violated? If so, which
ones?
```{r}
result <- lm(brain~body, data=mammals)
yhat = result$fitted.values
res = result$residuals
data = data.frame(mammals, yhat, res)
data %>%
  ggplot(aes(yhat, res)) + geom_point() + theme_bw() + geom_hline(yintercept=0, color="red")
```

# 3 
Based on your answers to parts 1 and 2, do we need to transform at least one of the
variables? Briefly explain.

# 4
For the simple linear regression in part 2, create a Box Cox plot. What transformation,
if any, would you apply to the response variable? Briefly explain.
y = y^0.1
```{r}
result <- lm(brain~body, data=mammals)
boxcox(result)
boxcox(result, lambda = seq(-0.05, 0.2, 1/100))
```
# 5
Apply the transformation you specified in part 4, and let y∗ denote the transformed
response variable. Create a scatterplot of y∗ against x. Comment on the appearance
of the plot. Do any assumptions for simple linear regression appear to be violated? If
so, which ones?

```{r}
temp <- mammals
temp$brain = temp$brain^0.075
temp %>%
  ggplot(aes(body, brain)) + geom_point() + theme_bw()
```

# 6
Fit a simple linear regression to y∗ against x, and create the corresponding residual
plot. Do any assumptions for simple linear regression appear to be violated? If so,
which ones?

```{r}
result <- lm(brain~body, data=temp)
yhat = result$fitted.values
res = result$residuals
data = data.frame(mammals, yhat, res)
data %>%
  ggplot(aes(yhat, res)) + geom_point() + theme_bw() + geom_hline(yintercept=0, color="red")
```

# 7
Do we need to transform the x variable? If yes, what transformation(s) would you try?
Briefly explain. Create a scatterplot of y∗ against x∗. Do any assumptions for simple
linear regression appear to be violated? If so, which ones?

```{r}
temp <- mammals
temp$brain = temp$brain^0.075
temp %>%
  ggplot(aes(log(body), brain)) + geom_point() + theme_bw()
```

# 8
Fit a simple linear regression to y∗ against x∗, and create the corresponding residual
plot. Do any assumptions for simple linear regression appear to be violated? If so,
which ones? If the assumptions are not met, repeat with a different transformation on
the predictor until you are satisfied.

```{r}
result <- lm(brain~log(body), data=temp)
yhat = result$fitted.values
res = result$residuals
data = data.frame(mammals, yhat, res)
data %>%
  ggplot(aes(yhat, res)) + geom_point() + theme_bw() + geom_hline(yintercept=0, color="red")

```

# 9
Create an ACF plot of the residuals. Comment if assumptions are met for linear
regression.
```{r}
acf(result$residuals, main="ACF Plot of Residuals with ystar")
```


# 10
Create a QQ plot of the residuals. Comment if assumptions are met for linear regression.
```{r}
qqnorm(result$residuals)
```


# 11
Write out the regression equation, and if possible, interpret the slope of the regression.
```{r}
result
```

$y^{0.075} = 1.00008012 + 0.002822(log(x))$
